---
title: "Parallelised computing as a method for increased performance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallelised computing as a method for increased performance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(duflor.gui)
```

# Introduction

Executing the analysis can takes varying amounts of time. There are 4 factors which influence the required time of analysis in a linearly proportional manner:

-   Number of checked pixels
-   Number of checked images
-   Number of spectra checked-for.
-   Image dimensions

For each spectrum one wants to identify, every pixel must be checked to fall within or outside the respectively defined HSV-boundaries.

It follows that in general, the following calculation yields the total number of iterations required to calculate the results:

$\text{Iterations}_{total} = \text{Checked pixels}_\text{total} = N_{Images}\times({Width}_{Image}\times{Height}_{Image})\times N_{spectra}$

This step is not particularly complicated, nor is it particularly resource-intensive. However,the sheer volume of iterations make this operation rather slow.

Additionally, the process of *loading* an image into an R-object to work on seems to be much more time-consuming than the analysis described above itself. Unfortunately, this steps is cannot be performed any faster currently. It is a bottle-neck that lies outside the scope of the project.

## The general steps

The following steps **must be performed** to obtain the results for a single image:

1.  Load the image into an R-object
2.  For each declared spectrum, check all pixels of the image against the respective upper and lower bounds, and save the coordinates of pixels fulfilling the condition
3.  Insert the results into an object persistent across spectrum-iterations
4.  Convert all results into a results-object

If multiple images are evaluated, the steps above must be performed the corresponding number of times.

## A reasonable worst-case scenario

Let's model a reasonable likely worst-case scenario:

-   250 images
-   2 spectra
-   Image-width: 6000 pixels
-   Image-height: 4000 pixels

To obtain all results, 250 images must be loaded. Without parallelisation (i.e. in sequential execution mode), every subsequent step requires its previous step to have concluded. Using the formula above, a total of

$$
\text{Iterations}_{total} = \text{Checked pixels}_\text{total} = N_{Images}\times({Width}_{Image}\times{Height}_{Image})\times N_{spectra}\\
 = 250\times(6000\times4000)\times2=1.2\times 10^{10}
$$

comparisons between a pixel's HSV-triplet and a lower- and upper-bound's HSV-triplet must be performed. Note: the above calculation is for a hypothetical root-area analysis of 250 images, where two spectra must be quantified (roots and the identifier-area itself).

While a single such check is performed basically instantaneously, it still takes *some* time.

# How a parallelised setup can help

R itself is a single-threaded language, and thus will utilise exactly *one* thread of *one* CPU-core to perform the iterations calculated above - by default.\
Thankfully, R has packages which allow the parallelised use of multiple available CPU-cores to work on operations simultaneously.

Thus, this app allows the user to run the entire analysis-pipeline in parallel, leveraging the `{foreach}`-package. It should be stressed that the *entire* pipeline is parallelised, **including** the loading-step.

## Advantages

Speed. By processing multiple images in parallel, the total time can be (roughly) divided by the number of available workers.

## Limitations

-   Setup-time: The process of setting up the parallel back-end takes a certain amount of time[^1]
-   RAM-usage: The amount of required RAM scales based on the number of used workers, the size of the images which are analysed, and the operating system.
-   Operating system: Without getting into the low-level details, there is a significant difference between "Windows"-systems and "Linux"/"MacOS" when it comes to managing back-ends for parallel-computing. Apple- and Linux-machines have access to the so-called `fork`-back-end. All three have access to the `socket`-back-end.

[^1]: Typically on the scale of single-digit seconds

### back-ends: `socket` vs `fork`

A so-called **`socket`**-back-end can be run on Windows, MacOS and Linux.\
It works by separately exporting the required environment (variables, functions, hidden objects, loaded packages, etc.) from the master-process to every designated worker, which work secluded from each other.\
When a worker finishes its workload, it must signal back to the master-process that it has finished. Once all workers are done, their results can be collected by the master, combined into a collective output format, and used as a "normal" variable from there.

The major disadvantage is the significant overhead accumulating by this back-and-forth, and thus a `socket`-cluster is **slow** (compared to the `fork`-back-end described below). For the reasons described above, a `socket`-back-end also heavily increases RAM-usage (linearly with the number of workers).

On the other hand, the **`fork`**-back-end runs significantly faster.\
Instead of duplicating all data and providing a unique copy to each worker-process, each worker gains access to the master's environment. As a consequence, the significant overhead of duplicating the environment can be ignored, and workers instead just return their respective results to the master process.\
Unfortunately, **this back-end is not available on windows**. As an additional side note, machine-clusters cannot use `fork`-back-ends.\
\

## Suggestions

When setting up a parallel back-end via the app, you should consider the specifications of your machine as described above:

1.  You might be tempted to create a parallel back.end using all available cores of a system. **DO NOT DO SO.** Simply don't. Assigning all available cores to your cluster will leave no core freely-available for other programs on the machine to do their job. This can and most certainly **will** cause havoc. Thus, leave at least a single core available. When setting up parallelisation via the shiny-GUI, it is not possible to use all cores available.
2.  If you have a limited amount of available RAM, there comes a point at which adding additional workers will deteriorate the performance of the entire back-end. If your machine has e.g. 8 GB of available RAM, 2 workers might work efficiently, and 3 might work just as fast as a sequential back-end[^2]. There is no clear guide to determine this tipping-point. In general, if your cluster comes close to consuming all available system RAM, adding more workers is likely to deteriorate overall performance, and slightly reducing the number of workers might even be beneficial.\

[^2]: In fact, 3 might be as fast as a non-parallelised back-end, or might even be slower. The same goes for adding even more workers.
