---
title: "How to report issues, raise suggestions and feedback"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to report issues, raise suggestions and feedback}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(duflor.gui)
```

This article mostly deals with using 


When issues arise, they **must be tracked on the package's GitHub page**, under "`r utils::packageDescription("duflor.gui")$BugReports`".

# General etiquette and different types of 'issues'

Although they are all conglomerated under the same name "issues", generally reports are differentiated into two categories:

- Bug Reports
- Feature Requests/Suggestions

## Bug Reports

Bug reports... are somewhat obvious at first. However, there is a great difference between what a good, a bad, and a useless bug-report.

To be useful, there is a bare-minimum amount of information that must be provided, including

- The package-versions of `duflor` and `duflor.gui` which are used.
   - If the package is run from source and is under version-control via `git` on the local machine (instead of being an installed package within R), the SHA of the currently checked-out commit must be provided.
- Detailed, clear instructions required to reproduce the issue.
- An example-dataset which exhibits the problem. In the case `duflor.gui` and `duflor`, this would often include example-images.
- Information on the system's specifications (OS, hardware configuration, etc. pp.)
- The R-Version in use
- notes on all warnings/errors emitted to the console before/when encountering the bug.


The more information is available, the easier it becomes to identify the actual issue.


## Feature Requests

To keep development efficient and reasonable, there are a certain number of requirements for reporting problems or requesting features.
A feature request should clearly define what additional feature should be implemented. As with bug-reports, the more information provided, the easier it's viability and ease of implementation can be assessed.

It should clearly describe _what_ should be implemented, as well as _why_ (and in broad strokes _how_) it should be implemented.
